Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-04-27 ThesisJP/Paper3/moving_obj.tex

\begin{abstract}
%\boldmath
If an ultrasound imaging system provides a presentation of a moving object which is sensitive to small spatial shifts, the system is said to be locally spatially shift-variant. This can for instance happen if the axial or lateral sampling is insufficient. The Capon beamformer has been shown to provide increased lateral resolution in ultrasound images. Increased lateral resolution should demand denser lateral sampling. However, in previous literature on Capon beamforming for medical ultrasound imaging only single frame scenarios have been simulated. Temporal behaviour and effects caused by the increased resolution and lack of oversampling have therefore been neglected.

In this paper we analyse the local lateral shift-invariance of the Capon beamformer when imaging moving objects. We show that insufficient lateral sampling makes an imaging system based on the Capon beamformer laterally shift-variant. Different methods for oversampling on transmit and receive are then discussed and investigated in order to improve on it. It is shown that lateral shift-invariance can be improved by oversampling based on phase rotation on receive. This without affecting the acquisition frame rate and with a minor change in processing complexity.
\end{abstract}

@@

In the last decade several authors have applied the Capon beamformer [also known as the minimum variance beamformer] for medical ultrasound imaging \cite{Synnevag2007, Vignon2008, Viola}. The method has been shown to improve lateral resolution and edge sharpness in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. This is accomplished by adaptively adjusting the array weights based on element-to-element covariance in order to minimize interference. The method is, however, also highly sensitive to model errors \cite{Mestre2006, Widrow1982, Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying steering vector uncertainty sets \cite{Lorenz2005, Rubsamen2013}. Spatial smoothing and diagonal loading, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics for high resolution parameters \cite{Synnevag2007a}. Steering vector uncertainty sets, however, have not been applied for medical ultrasound imaging.

@@

A steering vector uncertainty set reduces signal-nulling which appears if the signal of interest is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the Capon beamformer obtains its super resolution capabilities, but the signal can also get lost in the process. In \cite{JianLi2003} it was shown that most methods based on steering vector uncertainty sets are equivalent to diagonal loading where the loading factor depends on the steering vector uncertainty. In that way they make up for a given uncertainty by making the Capon beam wider. Hence, the maximal achievable resolution is decreased if the initial steering vector spacing was too coarse. In medical ultrasound imaging this initial set of steering vectors is governed by the number of receive beams. In this paper we will make use of an additional set of steering vectors to reduce the self-nulling effect by means of oversampling. The maximal achievable resolution is then maintained without the presence of severe self-nulling.

In \cite{Cox1973}, Cox presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. He also states the following important fact about the Capon beamformer [the $\vec{k}_3$ processor]: 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to four times the array resolution using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the high precision scanned output when the object is located at plus and minus half the required beam spacing for DAS [See (\ref{eq:resolution})]. %No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. 
The effect of typical subarray averaging and diagonal loading have been added to the Capon beamformer output [subarray length of 32 and a diagonal loading factor of 0.01]. We see that a DAS beamformer, using the required beam spacing, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However, if the Capon beamformer is used with the same beam density the loss could be as large as $27$ dB. This imaging system is then highly sensitive to small spatial shifts, either by the object or the array itself. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam spacing or object distance is decreased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=0.8\linewidth]{\img fig1_2way.eps}
}
\caption{Estimated drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming [with subarray length of 32 and a diagonal loading factor of 0.01] applied in a farfield-narrowband setting. %The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the required beam spacing for delay-and-sum, and for the latter the object is moved an angular distance corresponding to the system resolution. If the beam density was reduced to one beam per system resolution the maximum scalloping loss is found at zero degrees.
}
\label{fig:das_capon_beams}
\end{figure}

In previous literature on Capon beamforming for medical ultrasound imaging only single frame scenarios have been simulated. The scalloping loss effect has therefore been passed over in silence by either oversampling on transmit, by carefully positioning point scatterers exactly on transmit-receive beams, or by ignorance. In our first work where Capon beamforming was applied on a loop of images containing moving objects \cite{Asen2012, Asen} we commented on the effect, but no solution was presented other than oversampling on transmit when simulating. The same was done by Jensen and Austeng \cite{Jensen2012}.

In this paper we investigate the issue further. We show that when a moving point scatterer is imaged with the Capon beamformer with the same beam spacing which is sufficient for DAS, the predicted scalloping loss is observed both in simulations and \textit{in vitro} data. The question is then, what can we do in order to improve on the situation without affecting the acquisition frame rate and processing throughout. %We try two answer this by studying the scalloping loss of the Capon beamforming and investigating different methods aimed at reducing it. We also discuss how the different methods will affect the acquisition frame rate and the processing complexity.

The next section gives an introduction to Capon beamforming together with a short introduction to lateral sampling and shift-invariance. In Section \ref{sec:capon_LLSI} we study the scalloping loss yield by the Capon beamformer and its local lateral shift-invariance property.  Different methods for reduced scalloping loss and by that improved local lateral shift-invariance are presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res_paper3} we present the results of applying the most promising method on \textit{in vitro} data. Finally we discuss our findings in Section \ref{sec:dis_paper3} and give our conclusions in Section \ref{sec:con_paper3}.

@@

\section{Background}
\subsection{Standard Beamforming}
The standard form of time domain array beamforming is the DAS beamformer where the output $z$ is calculated as:
\begin{align}\label{eq:das}
z[n] = \sum_{m = 0}^{M-1}w_m^*x_m[n - \Delta_m[n]] = \vec{w}^H\vec{x}[n],
\end{align}
where $M$ is the number of elements or channels in the array, $\Delta_m[n]$ is the per-element focusing and steering delay, and $\vec{w}$ is a weight vector or window function. The weight vector is typically selected in order to trade resolution for a lower side lobe level. Hence, the weights are typically real and their K-space response is then symmetric.

@@

The sample covariance matrix can be estimated for an active broadband system in the following way \cite{Synnevag2009}:
\begin{align}
\mat{\breve{R}}[n] = \frac{1}{N_LN_K}\sum_{n'=n-K}^{n+K} \sum_{l=0}^{N_L-1} \vec{x}_l[n']\vec{x}_l[n']^H,
\end{align}
where $K$ is selected proportional to the pulse length, $N_K = 2K + 1$, $N_L = M-L+1$, and $\vec{x}_l$ is a $L$ long subarray $[x_l[n], \dotso x_{l+L}[n]]$. Finally the matrix is loaded with a diagonal factor for numerical stability and increased robustness, 
\begin{align}
\epsilon &= d*\text{trace}\{\mat{\breve{R}}\}/L\\
\mat{\hat{R}} &= \mat{\breve{R}} + \epsilon\mat{I}.
\end{align}

@@

The required lateral sampling in an ultrasound image for a given center frequency, $f_c$, and bandwidth, $B$, is given by the aperture size, $D$, and the wavelength of the center frequency, $\lambda_c = c/f_c$, where $c$ is the speed of sound. The required Nyquist beam spacing is then:
\begin{align}
\delta\theta < \frac{\lambda_c}{2qD}, \label{eq:resolution}
\end{align}
where the factor two is the effect of a two-way acquisition \cite{Hergum2009}, and where $q$ is an oversampling factor which can be applied to reduce folding of frequencies between $f_c$ and $f_c + B/2$. In this paper we will use the $q$-factor to perform oversampling prior to Capon beamforming.

Equation (\ref{eq:resolution}) was used to calculate the object positions in Fig.\,\ref{fig:das_capon_beams}. Therefore, if (\ref{eq:resolution}) is not fulfilled, the resulting scalloping loss for the DAS beamformer could become visible in the image. The system output will then be sensitive to small spatial shifts. We say that the system is locally spatially shift-variant. Since Capon beamforming improves the lateral resolution we will focus on sensitivity to small lateral shifts, or in other words the local lateral shift-invariance property of the imaging system.%This invariance to local movements is what makes ultrasound imaging feasible in general, but it is specially important for quantitative techniques, like speckle tracking of tissue and blood, that correlate time-varying features. %In this paper we analyze the lateral shift-invariance property of the Capon beamformer by studying the beamformer output when applied on simulated and \textit{in vitro} images of moving objects.

In an ultrasound image with 50 dB dynamic range mapped to 256 gray levels, a 1 dB loss corresponds to 5 gray levels. This is approximately equal to the visibility threshold [Weber fraction of 2\%] for gray scale images. A loss larger than 1 dB could therefore end up being visible to the observer. %However, what eventually matters is whether the loss is visible in the image or not.%In our further analysis a system would be regarded as locally laterally shift-invariant if the scalloping loss is less than 1 dB.  %However, depending on viewing light conditions and e.g. the selected gray level mapping, a larger loss might be acceptable.

In order to visualize the local lateral shift-invariance of a given imaging system, we will make use of lateral shift-variance plots (LSV-plots) \cite{Hergum2007}. A LSV-plot is constructed for a given imaging system by imaging a point scatterer moving laterally at constant speed. For each lateral position the root-mean-square (RMS) of the image data is calculated in range, and the resulting beam profiles are stacked in to an image. %For a given scan sequence there are twice as many point positions than there are receive lines. 
The LSV-plots in this paper are displayed as a contour plot with contours on -1, -2, -3, -6, -12 and -24 dB. The maximum for each beam profile is also marked with a black dot, and the amplitude variation among these points is presented in a subplot next to the LSV-plot. The LSV-plots based on simulations have been up-interpolated to 128 samples in each direction using cubic interpolation if the number of receive beams is less than 128.

To further quantify the shift-invariance we have calculated the mean absolute error (MAE) between the maximum point position in each beam profile and a straight line ($f(x)=x$). We will refer to this measure as P-MAE. The MAE of the maximum points' amplitude versus the data maximum is also measured. We refer to this measure as A-MAE. The first measure quantifies the amount of geometric distortion \cite{Hergum2007}, and the latter quantifies the peak gain variation or in other words the scalloping loss. The maximum amplitude deviation, or maximum scalloping loss, along this point trace is also calculated. We present all these numbers together with the LSV-plot.

@@

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). Dotted vertical lines indicate position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} using a 96 element phased array with a 22 mm azimuth aperture, which was excited with a 2.5 MHz and 1.5 cycles long pulse. Transmit focus in azimuth and elevation were configured to be equal and the point scatterer was moved through this point. The transmit focus is where the ultrasound beam is at its narrowest, hence shift variance effects will appear at this location first. The beam density was calculated  based on (\ref{eq:resolution}) with a speed of sound equal to 1540. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. %In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located in the center of this beam at one time instance.

@@

We observe that Fig.\,\ref{fig:capon2} now contains a scaled version of the pattern presented in Fig.\,\ref{fig:das_shiftvariant}. From Fig.\,\ref{fig:capon} it should be clear that Capon beamforming either requires increased lateral sampling, or the resolution has to be reduced to the same level as DAS in order to make the imaging system locally laterally shift-invariant. It would be unwise to do all the computation involved with Capon beamforming with a high $d$ value and end up with a resolution equal to DAS. However, we see that diagonal loading can be a useful tool if resolution has to be traded for the need of oversampling. The same is true for the subarray length parameter.

@@

\section{Oversampling Methods}\label{sec:methods}
In this section we will discuss different methods aimed at improving the local lateral shift-invariance of the Capon beamformer. From the previous section it should be clear that we somehow need to increase the lateral sampling to make the Capon imaging system locally laterally shift-invariant. In this section we will investigate how this oversampling should be conducted, how large it needs to be and how the proposed methods will affect the acquisition frame rate and the processing complexity.

\subsection{Oversampling on Transmit}
A straight forward way of obtaining sufficient sampling in an active imaging system with an equal number of transmit (tx) and receive (rx) beams is to increasing the number of tx-beams until the system is locally laterally shift-invariant. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the tx-beam density by a $q$-factor of four and sixteen respectively. As expected the geometric distortion and scalloping loss decreases with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed compared to Fig.\,\ref{fig:capon} in order to better see the beam-to-beam variation.

From Fig.\,\ref{fig:capon_oversampling_b} we see that sixteen times oversampling reduces the scalloping loss significantly. Even though high frequency variations are reduced, the maximum is still larger than 1 dB. An even higher level of oversampling is therefore required in order to have a completely local lateral shift-invariant Capon beamformer when a moving point scatterer is simulated without noise and with the same parameters as in Fig.\,\ref{fig:capon}. %$1/100$ in diagonal loading. 
Note that a diagonal loading factor of $1/100$ will introduce a white noise component in the sample covariance matrix $20$ dB lower than the signal present. Hence, the covariance has been estimated from data which appeared to have an SNR of 20 dB.

Even though increasing the number of transmits improves the lateral shift-invariance it will also reduce the acquisition frame rate with a factor equal to the oversampling factor. The processing complexity will also increase equally. For these two reasons, oversampling on transmit is something which is desirable to avoid in real time ultrasound applications, and therefore we will not discuss this method further.

\begin{figure*}[!t]
\centerline{
\begin{tabular}{c}
\subfloat[]{
\includegraphics[width=0.7\linewidth]{\img fig4a.eps}\label{fig:capon_oversampling_a}
}\\
\subfloat[]{
\includegraphics[width=0.7\linewidth]{\img fig4b.eps}\label{fig:capon_oversampling_b}
}
\end{tabular}
}
\caption{Capon beamforming as described in Fig.\,\ref{fig:capon} with oversampling on transmit. The figure is zoomed compared with Fig.\,\ref{fig:capon} to better see the beam-to-beam variation. a) Four times oversampling ($q=4$). b) Sixteen times oversampling ($q=16$). The maximum scalloping loss is reduced, but it is still too high for the system to be locally laterally shift invariant.}
\label{fig:capon_oversampling}
\end{figure*}

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams in order to get sufficient sampling on receive, it is possible to achieve high frame rate imaging. However, PRB also introduces geometrical distortion which to some degree can be compensated for by using e.g. synthetic transmit beamforming (STB) \cite{Hergum2007, Denarie2013}. In \cite{Rabinovich2013} it is shown that STB works well with Capon beamforming, something which is also verified in \cite{Asen}. In the following discussion we consider corrected PRB as being equal to oversampling on transmit when it comes to image quality. The LSV-plots for PRB are therefore not included.

In cardiac ultrasound imaging, PRB is typically used in order to keep the number of rx-beams high while the number of tx-beams are reduced. When using PRB for oversampling we would like to maintain the same number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand, more rx-beams will increase the number of covariance matrices and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with the same factor. Since more than sixteen times oversampling is required in order to achieve lateral-shift-invariant and high-resolution imaging of a moving point scatterer, none of the implementations described in \cite{Asen} would be real time any more. We therefore need a different way of oversampling our imaging sector.

>>>>

\begin{abstract}
%\boldmath
If an ultrasound imaging system provides a presentation of a moving object which is sensitive to small spatial shifts, the system is said to be locally spatially shift-variant. This can, for instance, happen if the axial or lateral sampling is insufficient. The Capon beamformer has been shown to provide increased lateral resolution in ultrasound images. Increased lateral resolution should demand denser lateral sampling. However, in previous literature on Capon beamforming for medical ultrasound imaging only single-frame scenarios have been simulated. Temporal behaviour and effects caused by the increased resolution and lack of oversampling have therefore been neglected.

In this paper, we analyse the local lateral shift-invariance of the Capon beamformer when imaging moving objects. We show that insufficient lateral sampling makes an imaging system based on the Capon beamformer laterally shift-variant. Different methods for oversampling on transmit and receive are then discussed and investigated to improve on the Capon beamformer. It is shown that lateral shift-invariance can be improved by oversampling based on phase rotation on receive. This without affecting the acquisition frame rate and with a minor change in processing complexity.
\end{abstract}

@@

In the last decade several authors have applied the Capon beamformer [also known as the minimum variance beamformer] for medical ultrasound imaging \cite{Synnevag2007, Vignon2008, Viola}. The method has been shown to improve lateral resolution and edge sharpness in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. This is accomplished by adaptively adjusting the array weights based on element-to-element covariance to minimize interference. The method is, however, also highly sensitive to model errors \cite{Mestre2006, Widrow1982, Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g., spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or application of steering vector uncertainty sets \cite{Lorenz2005, Rubsamen2013}. Spatial smoothing and diagonal loading, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics for high resolution parameters \cite{Synnevag2007a}. Steering vector uncertainty sets, however, have not been applied for medical ultrasound imaging.

@@

A steering vector uncertainty set reduces signal-nulling which appears if the signal of interest is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the Capon beamformer obtains its super resolution capabilities, but the signal can also get lost in the process. In \cite{JianLi2003}, it was shown that most methods based on steering vector uncertainty sets are equivalent to diagonal loading where the loading factor depends on the steering vector uncertainty. In that way, they make up for a given uncertainty by making the Capon beam wider. Hence, the maximal achievable resolution is decreased if the initial steering vector spacing was too coarse. In medical ultrasound imaging, this initial set of steering vectors is governed by the number of receive beams. In this paper, we will make use of an additional set of steering vectors to reduce the self-nulling effect by means of oversampling. The maximal achievable resolution is then maintained without the presence of severe self-nulling.

In \cite{Cox1973}, Cox presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. He also states the following important fact about the Capon beamformer (the $\vec{k}_3$ processor): 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using \cite[Eqs. (26) and (32)]{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to four times the array resolution using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the high precision scanned output when the object is located at plus and minus half the required beam spacing for DAS [See (\ref{eq:resolution})]. %No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. 
The effect of typical subarray averaging and diagonal loading have been added to the Capon beamformer output [subarray length of 32 and a diagonal loading factor of 0.01]. We see that a DAS beamformer, using the required beam spacing, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However, if the Capon beamformer is used with the same beam density, the loss could be as large as $27$ dB. This imaging system is then highly sensitive to small spatial shifts, either by the object or the array itself. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam spacing or object distance is decreased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=0.8\linewidth]{\img fig1_2way.eps}
}
\caption{Estimated drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming (with subarray length of 32 and a diagonal loading factor of 0.01) applied in a far-field-narrowband setting. %The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the required beam spacing for delay-and-sum, and for the latter the object is moved an angular distance corresponding to the system resolution. If the beam density was reduced to one beam per system resolution the maximum scalloping loss is found at zero degrees.
}
\label{fig:das_capon_beams}
\end{figure}

In previous literature on Capon beamforming for medical ultrasound imaging, only single frame scenarios have been simulated. The scalloping loss effect has therefore been passed over in silence by either oversampling on transmit, by carefully positioning point scatterers exactly on transmit-receive beams, or by ignorance. In our first work in which Capon beamforming was applied on a loop of images containing moving objects \cite{Asen2012, Asen}, we commented on the effect, but no solution was presented other than oversampling on transmit when simulating. The same was done by Jensen and Austeng \cite{Jensen2012}.

In this paper, we investigate the issue further. We show that when a moving point scatterer is imaged with the Capon beamformer with the same beam spacing which is sufficient for DAS, the predicted scalloping loss is observed both in simulations and \textit{in vitro} data. The problem is then to find what can be done to improve on the situation without affecting the acquisition frame rate and processing throughout. %We try two answer this by studying the scalloping loss of the Capon beamforming and investigating different methods aimed at reducing it. We also discuss how the different methods will affect the acquisition frame rate and the processing complexity.

The next section gives an introduction to Capon beamforming together with a short introduction to lateral sampling and shift-invariance. In Section \ref{sec:capon_LLSI} we study the scalloping loss yield by the Capon beamformer and its local lateral shift-invariance property.  Different methods for reduced scalloping loss and by that improved local lateral shift-invariance are presented in Section \ref{sec:methods}. Then, in Section \ref{sec:res_paper3}, we present the results of applying the most promising method on \textit{in vitro} data. Finally, we discuss our findings in Section \ref{sec:dis_paper3} and give our conclusions in Section \ref{sec:con_paper3}.

@@

\section{Background}
\subsection{Standard Beamforming}
The standard form of time-domain array beamforming is the DAS beamformer, for which the output $z$ is calculated as
\begin{align}\label{eq:das}
z[n] = \sum_{m = 0}^{M-1}w_m^*x_m[n - \Delta_m[n]] = \vec{w}^H\vec{x}[n],
\end{align}
where $M$ is the number of elements or channels in the array, $\Delta_m[n]$ is the per-element focusing and steering delay, and $\vec{w}$ is a weight vector or window function. The weight vector is typically selected to trade resolution for a lower side-lobe level. Hence, the weights are typically real and their K-space response is then symmetric.

@@

The sample covariance matrix can be estimated for an active broadband system in the following way \cite{Synnevag2009}:
\begin{align}
\mat{\breve{R}}[n] = \frac{1}{N_LN_K}\sum_{n'=n-K}^{n+K} \sum_{l=0}^{N_L-1} \vec{x}_l[n']\vec{x}_l[n']^H,
\end{align}
where $K$ is selected proportional to the pulse length, $N_K = 2K + 1$, $N_L = M-L+1$, and $\vec{x}_l$ is a $L$-long subarray $[x_l[n], \dotso x_{l+L}[n]]$. Finally the matrix is loaded with a diagonal factor for numerical stability and increased robustness, 
\begin{align}
\epsilon &= d*\text{trace}\{\mat{\breve{R}}\}/L\\
\mat{\hat{R}} &= \mat{\breve{R}} + \epsilon\mat{I}.
\end{align}

@@

The required lateral sampling in an ultrasound image for a given center frequency, $f_c$, and bandwidth, $B$, is given by the aperture size, $D$, and the wavelength of the center frequency, $\lambda_c = c/f_c$, where $c$ is the speed of sound. The required Nyquist beam spacing is then:
\begin{align}
\delta\theta < \frac{\lambda_c}{2qD}, \label{eq:resolution}
\end{align}
where the factor two is the effect of a two-way acquisition \cite{Hergum2009}, and where $q$ is an oversampling factor which can be applied to reduce folding of frequencies between $f_c$ and $f_c + B/2$. In this paper, we will use the $q$-factor to perform oversampling before Capon beamforming.

Equation (\ref{eq:resolution}) was used to calculate the object positions in Fig.\,\ref{fig:das_capon_beams}. Therefore, if (\ref{eq:resolution}) is not fulfilled, the resulting scalloping loss for the DAS beamformer could become visible in the image. The system output will then be sensitive to small spatial shifts. We say that the system is locally spatially shift-variant. Since Capon beamforming improves the lateral resolution we will focus on sensitivity to small lateral shifts, or in other words, the local lateral shift-invariance property of the imaging system.%This invariance to local movements is what makes ultrasound imaging feasible in general, but it is specially important for quantitative techniques, like speckle tracking of tissue and blood, that correlate time-varying features. %In this paper we analyze the lateral shift-invariance property of the Capon beamformer by studying the beamformer output when applied on simulated and \textit{in vitro} images of moving objects.

In an ultrasound image with 50 dB dynamic range mapped to 256 gray levels, a 1 dB loss corresponds to 5 gray levels. This is approximately equal to the visibility threshold [Weber fraction of 2\%] for grayscale images. A loss larger than 1 dB could therefore end up being visible to the observer. %However, what eventually matters is whether the loss is visible in the image or not.%In our further analysis a system would be regarded as locally laterally shift-invariant if the scalloping loss is less than 1 dB.  %However, depending on viewing light conditions and e.g. the selected gray level mapping, a larger loss might be acceptable.

To visualize the local lateral shift-invariance of a given imaging system, we will make use of lateral shift-variance plots (LSV-plots) \cite{Hergum2007}. An LSV-plot is constructed for a given imaging system by imaging a point scatterer moving laterally at constant speed. For each lateral position, the rms of the image data is calculated in range, and the resulting beam profiles are stacked in to an image. %For a given scan sequence there are twice as many point positions than there are receive lines. 
The LSV-plots in this paper are displayed as a contour plot with contours on -1, -2, -3, -6, -12 and -24 dB. The maximum for each beam profile is also marked with a black dot, and the amplitude variation among these points is presented in a subplot next to the LSV-plot. The LSV-plots based on simulations have been up-interpolated to 128 samples in each direction using cubic interpolation if the number of receive beams is less than 128.

To further quantify the shift-invariance we have calculated the mean absolute error (MAE) between the maximum point position in each beam profile and a straight line [$f(x)=x$]. We will refer to this measure as P-MAE. The MAE of the maximum points' amplitude versus the data maximum is also measured. We refer to this measure as A-MAE. The first measure quantifies the amount of geometric distortion \cite{Hergum2007}; and the latter quantifies the peak gain variation, or in other words, the scalloping loss. The maximum amplitude deviation, or maximum scalloping loss, along this point trace is also calculated. We present all these numbers together with the LSV-plot.

@@

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). Dotted vertical lines indicate position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} using a 96 element phased array with a 22 mm azimuth aperture, which was excited with a 2.5 MHz and 1.5 cycles long pulse. Transmit focus in azimuth and elevation were configured to be equal and the point scatterer was moved through this point. The transmit focus is where the ultrasound beam is at its narrowest; hence shift variance effects will appear at this location first. The beam density was calculated  based on (\ref{eq:resolution}) with a speed of sound equal to 1540. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. %In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located in the center of this beam at one time instance.

@@

We observe that Fig.\,\ref{fig:capon2} now contains a scaled version of the pattern presented in Fig.\,\ref{fig:das_shiftvariant}. From Fig.\,\ref{fig:capon} it should be clear that Capon beamforming either requires increased lateral sampling, or the resolution has to be reduced to the same level as DAS to make the imaging system locally laterally shift-invariant. It would be unwise to do all the computation involved with Capon beamforming with a high $d$ value and end up with a resolution equal to DAS. However, we see that diagonal loading can be a useful tool if resolution has to be traded for the need of oversampling. The same is true for the subarray length parameter.

@@

\section{Oversampling Methods}\label{sec:methods}
In this section, we will discuss different methods aimed at improving the local lateral shift-invariance of the Capon beamformer. From the previous section, it should be clear that we somehow need to increase the lateral sampling to make the Capon imaging system locally laterally shift-invariant. In this section, we will investigate how this oversampling should be conducted, how large it must be, and how the proposed methods will affect the acquisition frame rate and the processing complexity.

\subsection{Oversampling on Transmit}
A straightforward way of obtaining sufficient sampling in an active imaging system with an equal number of transmit (tx) and receive (rx) beams is to increasing the number of tx-beams until the system is locally laterally shift-invariant. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the tx-beam density by $q$-factors of 4 and 16 respectively. As expected, the geometric distortion and scalloping loss decrease with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed-in compared to Fig.\,\ref{fig:capon} to better see the beam-to-beam variation.

From Fig.\,\ref{fig:capon_oversampling_b}, we see that 16 times oversampling reduces the scalloping loss significantly. Even though high-frequency variations are reduced, the maximum is still larger than 1 dB. An even higher level of oversampling is therefore required to have a completely local lateral shift-invariant Capon beamformer when a moving point scatterer is simulated without noise and with the same parameters as in Fig.\,\ref{fig:capon}. %$1/100$ in diagonal loading. 
Note that a diagonal loading factor of $1/100$ will introduce a white noise component in the sample covariance matrix $20$ dB lower than the signal present. Hence, the covariance has been estimated from data which appeared to have an SNR of 20 dB.

Even though increasing the number of transmits improves the lateral shift-invariance, it will also reduce the acquisition frame rate with a factor equal to the oversampling factor. The processing complexity will also increase equally. For these two reasons, oversampling on transmit is something which is desirable to avoid in real time ultrasound applications, and therefore we will not discuss this method further.

\begin{figure*}[!t]
\centerline{
\begin{tabular}{c}
\subfloat[]{
\includegraphics[width=0.7\linewidth]{\img fig4a.eps}\label{fig:capon_oversampling_a}
}\\
\subfloat[]{
\includegraphics[width=0.7\linewidth]{\img fig4b.eps}\label{fig:capon_oversampling_b}
}
\end{tabular}
}
\caption{Capon beamforming as described in Fig.\,\ref{fig:capon} with oversampling on transmit. The figure is zoomed-in compared with Fig.\,\ref{fig:capon} to better see the beam-to-beam variation. a) 4 times oversampling ($q=4$). b) 16 times oversampling ($q=16$). The maximum scalloping loss is reduced, but it is still too high for the system to be locally laterally shift invariant.}
\label{fig:capon_oversampling}
\end{figure*}

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams to get sufficient sampling on receive, it is possible to achieve high frame rate imaging. However, PRB also introduces geometrical distortion which, to some degree, can be compensated for by using, e.g., synthetic transmit beamforming (STB) \cite{Hergum2007, Denarie2013}. In \cite{Rabinovich2013}, it is shown that STB works well with Capon beamforming, something which is also verified in \cite{Asen}. In the following discussion, we consider corrected PRB as being equal to oversampling on transmit when it comes to image quality. The LSV-plots for PRB are therefore not included.

In cardiac ultrasound imaging, PRB is typically used to keep the number of rx-beams high while the number of tx-beams are reduced. When using PRB for oversampling, we would like to maintain the same number of tx-beams while the number of rx-beams are increased. The acquisition frame rate will then not be affected. On the other hand, more rx-beams will increase the number of covariance matrices and thereby the number of matrix-inversions with a factor equal to the number of parallel receive lines. The processing times reported in \cite{Asen} will then also increase with the same factor. Because more than 16 times oversampling is required to achieve lateral-shift-invariant and high-resolution imaging of a moving point scatterer, none of the implementations described in \cite{Asen} would be possible in real time. We therefore need a different way of oversampling our imaging sector.

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ because ultrasound data must be pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of predefined vectors. In a narrow band setting, the beam can make a sweep across the whole imaging sector through phase steering. This is also possible in broadband applications as long as the phase rotation is less than one pulse length (See coarse-fine beamforming in \cite{Thomenius} and its references). The maximal steering angle is approximately given by
\begin{align}
\theta_{max} = 2\alpha \delta\theta,
\end{align}
where $\alpha$ is the pulse length in terms of $\lambda$, and $2$ and $\delta\theta$ is the oversampling factor and beam spacing from (\ref{eq:resolution}). For our simulations $\theta_{max} = 3\delta\theta$.
It is therefore possible to make a sweep based on phase rotation between a given beam and half-way to its two nearest neighbours without introducing large errors. Phase rotation is actually what the Capon beamformer uses internally to micro-steer away from bright point scatterers and to enhance edges.
